---
title: "Notebook"
---

These lines clear variables, plots and help pages:

```{r}
rm(list = ls())
dev.off()
help.start()
```

We first need to get the appropriate library ready. Remember to precede every line with the `install.packages` command.

```{r}
library("astsa")
library("xts")
library("readr")
library("lubridate")
library("dplyr")
```

We're going to extract the data from Arpa Lombardia.

```{r}
setwd("C:/Users/39349/OneDrive/Desktop/Projects/Su-git-hub/rain-chaos")
getwd()
df <- read.csv("Precipitazioni_dal_2011_al_2020_20250215.csv", header = TRUE)
```
This data frame contains rain precipitation measurements from 2014-07-18 to 2021-01-01, done by a single station every 10 minutes. We only care about `"Data"` and `"Valore"`, thus:
```{r}
df <- df[, c("Data", "Valore")]
```

Let's first focus on converting the dates into a proper format, instead of just strings.

```{r}
df$Data <- dmy_hms(df$Data)
```
It's easy to see that this data frame is not sorted date-wise, so we immediately fix that for time series plotting.

```{r}
is.unsorted(df$Data) # returns true
df <- df[order(df$Data), ] # reorders the data
```

The station had some periods of time in which it was not collecting any data, so we don't have the correct number of measurements, let's fix that.

```{r}
expected_times <- seq(from = min(df$Data), to = max(df$Data), by = "10 min")
missing_times <- setdiff(expected_times, df$Data)
# Check if there are missing timestamps
if (length(missing_times) > 0) {
  cat("There are", length(missing_times), "missing timestamps.\n")
} else {
  cat("No missing timestamps!\n")
}

# This inserts the missing times in the right order
df <- data.frame(Data = expected_times) %>% left_join(df, by = "Data")
# We get rid of the first measurement (I don't remember why I did that).
df <- df[df$Data >= "2014-07-19 00:00:00", ]
```

Now we can set up the time series

```{r}
xts_data <- xts(df$Valore, order.by = df$Data)
# This gets rid of faulty measurements. This could've been done earlier to be fair
xts_data[xts_data < 0] <- NA
```

For some reason (I don't remember) we take the sum of weekly measurements.

```{r}
weekly_meas <- apply.weekly(xts_data, sum)

weekly_meas <- weekly_meas[weekly_meas > 1] # Why are we only considering weekly measurements > 1?
summary(weekly_meas) # this reveals the quantiles
plot(weekly_meas)
```
```{r}

tau <- 1

weekly_meas_shifted <- lag.xts(weekly_meas, k = -1)

plot(as.numeric(weekly_meas), as.numeric(weekly_meas_shifted),
     xlab = "Weekly measurement (Current Week)", ylab = "Weekly measurement (Next Week)",
     main = "Weekly measurement vs Next Week's measurement", type = "l", col = "blue",
     lty = 1, lwd = 2)

```